---
title: "Approximating logistic coefficients"
subtitle: 'A simulation analysis'
author: "Developed by [Gabriel Hoffman](http://gabrielhoffman.github.io/)"
date: "Run on `r Sys.time()`"
documentclass: article
output: 
  html_document:
      fig_caption: yes
  toc: true
  smart: true
vignette: >
  %\VignetteIndexEntry{Approximating logistic coefficients}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\usepackage[utf8]{inputenc}
---


```{r setup, echo=FALSE, results="hide"}
knitr::opts_chunk$set(
  tidy = FALSE,
  cache = TRUE,
  echo = FALSE,
  dev = c("png", "pdf"),
  package.startup.message = FALSE,
  message = FALSE,
  error = FALSE,
  warning = FALSE
)
``` 

Given a z-statistic, we want to obtain the coefficient value from a linear regression. Adapting the approach from [Zhu, et al (2016, Methods eqn 6)](https://doi.org/10.1038/ng.3538), we estimate the coefficient as $$\beta = z * sd_y / (sd_x*sqrt(n + z^2)),$$ where \(sd_x\) is the standard deviation of the covariate and \(sd_y\) is the standard error of the response. For a model with no covariates, this transformation gives the exact coefficient estimate. With covariates, it is approximate.

Second, the coeffient estimate from linear regression can be converted to the logistic scale using the approach of [Pirinen, et al. (2013)](https://doi.org/10.1214/12-AOAS586) according to $$\beta / (\phi(1-\phi)),$$ where \(\phi\) is the case ratio in the logistic regression. This approximates the coefficient as if the model had been fit with logistic regression.  This is implemented in `coef_from_z()`.

We see here that using the z-statistic to approximate the coefficient from logistic regression model is most accurate for large sample size, balanced class ratio (i.e. $\phi$ near 0.5), and small $\beta$. 

```{r load, cache=FALSE}
library(imputez)
library(tidyverse)
library(ggplot2)
library(cowplot)

# simulate data

p = .5

get_estimates = function(mu, beta, n){

  x = rbinom(n, 2, p)
  x = x - mean(x)

  eta = mu + x*beta
  y = rbinom(n, size=1, prob=plogis(eta))
  data = data.frame(x, y)

  fit <- glm(y ~ x, data=data, family=binomial)
  z = coef(summary(fit))[2,3]
  phi = sum(y) / length(y)

  beta_logit = coef(summary(fit))[2,]

  beta_hat = coef_from_z(z, n, sd(x), phi=phi)

  data.frame(beta_logit = beta_logit[1], 
            beta_hat = beta_hat[1],
            phi = phi)
}
```

```{r sim1}
# Sample size
#############

n_array = 10^seq(2.2, 5, length.out=20)

res = lapply(floor(n_array), function(n){
  lapply(seq(20), function(i){
    set.seed(i)
    get_estimates(0, .1, n)
  }) %>%
  bind_rows %>%
  summarize(rMSE = sqrt(mean((beta_logit-coef)^2))) %>%
  mutate(n = n)
})
res = bind_rows(res)

thm = theme_classic() +
    theme(aspect.ratio=1)

fig1 = res %>%
  ggplot(aes(n, rMSE )) +
    geom_point() +
    thm +
    scale_x_log10() +
    scale_y_log10() +
    ylab("Root mean squared error")
```


```{r sim2}
# Phi
######

n = 1000
mu_array = seq(-2.5, 2.5, length.out=100)

res = lapply(mu_array, function(mu){
  lapply(seq(100), function(i){
    set.seed(i)
    get_estimates(mu, 0, n)
  }) %>%
  bind_rows %>%
  summarize(rMSE = sqrt(mean((beta_logit-coef)^2)), phi = mean(phi)) %>%
  mutate(mu = mu)
})
res = bind_rows(res)

fig2 = res %>%
  ggplot(aes(phi, rMSE )) +
    geom_point() +
    thm +
    scale_x_continuous(limits=c(0, 1), expand=c(0,0)) +
    xlab(bquote(phi)) +
    ylab("Root mean squared error")
```


```{r sim3}
# beta
######

n = 10000
mu = 0
beta_array = seq(-1, 1, length.out=100)

res = lapply(beta_array, function(beta){
  lapply(seq(10), function(i){
    set.seed(i)
    get_estimates(mu, beta, n)
  }) %>%
  bind_rows %>%
  summarize(rMSE = sqrt(mean((beta_logit-coef)^2))) %>%
  mutate(beta = beta)
})
res = bind_rows(res)

fig3 = res %>%
  ggplot(aes(beta, rMSE )) +
    geom_point() +
    thm +
    xlab(bquote(beta)) +
    ylab("Root mean squared error")
```


```{r plot, cache=FALSE, fig.width=10, fig.height=3.5}
plot_grid(fig1, fig2, fig3, nrow=1, labels=LETTERS[1:3])
```



# Session Info
<details>
```{r session, echo=FALSE}
sessionInfo()
```
</details>





